---
title: "casestudy 11"
output: 
  html_document:
    keep_md: TRUE
    code_folding: hide 
---

```{r setup, include=FALSE}
library(tidyquant)
library(tidyverse)
library(dplyr)
library(lubridate)
library(timetk)
library(dygraphs)

tickers <- c("CXW", "F", "GM", "JCP", "KR", "WDC", "NKE","T", "WDAY", "WFC", "WMT")
start <- today() - years(5)
prices <- tq_get(tickers, get = 'stock.prices', from = start)
head(prices)

matrix_prices <- prices %>% 
  select(symbol, date, adjusted) %>% 
  pivot_wider(names_from = symbol, values_from = adjusted) %>% 
  tk_xts(date_var = date)
head(matrix_prices)

perc_change <- prices %>% 
  group_by(symbol) %>% 
  mutate(percent_change_total = (adjusted/adjusted[1]-1)*100)
head(perc_change)

matrix_prices_perc <- perc_change %>% 
  select(symbol, date, percent_change_total) %>% 
  pivot_wider(names_from = symbol, values_from = percent_change_total) %>% 
  tk_xts(date_var = date)

volumes <- prices %>% 
  group_by(symbol, date, volume) %>% 
  select(symbol, date, volume)
head(volumes)

```

```{r}
dygraph(matrix_prices, main = 'Tickers') %>%
  dySeries(fillGraph = FALSE) %>% 
  dyHighlight(highlightCircleSize = 5., 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE) %>% 
  dyOptions(axisLineColor = "black",
            gridLineColor = "black") %>% 
  dyRangeSelector() 

```

```{r}
dygraph(matrix_prices, main = 'If $25,000 Were Invested Into Each 5 Years Ago') %>% 
  dySeries(fillGraph = FALSE) %>% 
  dyHighlight(highlightCircleSize = 5., 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE) %>%
  dyRebase(value = 25000)

```



```{r}
dygraph(matrix_prices_perc, main = 'Percent Change') %>%
  dySeries(fillGraph = FALSE) %>% 
  dyHighlight(highlightCircleSize = 5., 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE) %>% 
  dyOptions(axisLineColor = "black",
            gridLineColor = "black") %>% 
  dyRangeSelector()
```


```{r}
ggplot() +
  geom_line(data = volumes, mapping = aes(x = date, y = volume, color = symbol)) +
  facet_wrap(~symbol, nrow = 3) +
  labs(x = 'Date', y = 'Volume', title = 'Volumes For Each Stock') +
  theme_bw() +
  theme(legend.position = 'none')

```

## conclusion 

I made graphs of the stock prices, percent change from beginning price, the value with 25,000 dollars initially invested, and time series graph of volume. 

after viewing the graphs, I would invest in WMT, GM, WFC, NKE, and WDC. personally, I would invest 40% in WMT because it was the highest performing stock and rebounding well currently from the pandemic. the rest of the initial investment would be evenly distributed among the rest of the top perfoming stocks by percent change, GM, WFC, NKE, and WDC. 

as for volume, if a stock had a conceringly low stock price I would check the volume graph. if the volume was consistently high, I would infer that the stock is still healthy and just traded in higher volumes at a lower price, possibly due to stock splits. if the stock had a low price and volume, I would take it as a sign of an underperforming stock to avoid. 
