---
title: "casestudy 6"
output: 
  html_document:
    keep_md: TRUE
---

#1 
distributions of x,y, and z
```{r,echo=FALSE}
p5 <- ggplot(diamonds, aes(x = x)) +
  geom_density()
p5

p9 <- ggplot(diamonds, aes(x = y)) +
  geom_density()
p9

p10 <- ggplot(diamonds, aes(x = z)) + 
  geom_density()
p10
```

#2 
distribution of price
```{r,echo=FALSE}
p6 <- ggplot(diamonds, aes(x = price)) +
  geom_density(binwidth = 50, fill = 'light blue')
p6
```
Nothing too surprising, there is a heavy skew towards lower priced diamonds. The only possible abnormality would be the slight uptick of diamonds just below 5000 dollars. This could be due variation in the data though

#3
```{r,echo=FALSE}
p1 <- ggplot(diamonds, aes(x = price, y = x)) +
  geom_point() +
  geom_abline(slope = 1) +
  scale_x_continuous(trans = 'log10')
p1

p2 <- ggplot(diamonds, aes(x = price, y = y)) +
  geom_point() +
  geom_abline(slope = 1) +
  scale_x_continuous(trans = 'log10')
p2

p3 <- ggplot(diamonds, aes(x = price, y = z)) +
  geom_point() + 
  geom_abline(slope = 1) +
  scale_x_continuous(trans = 'log10')
p3
```
By the graphs we can see that z is the closest to the line denoting a 1:1 correlation and is therefore the best predictor of price. 
```{r,echo=FALSE}
p4 <- ggplot(diamonds, aes(x = cut, y = z)) +
  geom_violin() 
p4
```
As we can see from this graph, the trend of the data seems to travel direcetly opposite the 1:1 line, indicating a negative correlation of price and cut. This decreasing of cut quality as price increases shows how lower quality diamonds can be a higher value. This is probably due to the price of the diamonds being more determined by size than quality. 

#4
carat partitioned by price 
```{r,ehco=FALSE}
dat <- diamonds %>% 
  mutate(
    type = case_when(
      price < 2500 ~ '<2500',
      price < 5000 ~ '<5000',
      price < 7500 ~ '<7500',
      price < 10000 ~ '<10000',
      price < 12500 ~ '<12500',
      price < 15000 ~ '<15000',
      price < 17500 ~ '<17500',
      price < 20000 ~ '<20000',
      price < 22500 ~ '<22500',
      TRUE ~ '>22500')
  )

dat$type <- factor(dat$type, levels = str)  

p7 <- ggplot(dat, aes(x = type, y = carat)) +
  geom_violin()
p7
```

#5
price distriburion of very large vs very small diamonds 
```{r,echo=FALSE}
quantile(diamonds$carat)

dat2 <- diamonds %>% 
  mutate(
    type = case_when(
      carat <= .4 ~ 'very small',
      carat <=1.04 ~ 'normal',
      carat > 1.04 ~'very large'
    ) 
  ) %>% 
  filter(type != 'normal')
      


p8 <- ggplot(dat2, aes(x = type, y = price)) +
  geom_violin() 
p8
```
As we can see, the large diamonds have a very wide distribution, and the small diamonds have a very tight distribution concentrated at the low end 

#6
```{r}
require(lattice)
levelplot(diamonds$price ~ diamonds$cut*diamonds$carat,col.regions = rainbow(256), xlab = 'cut', ylab = 'carat')

```
The graph shows that across all cuts, the price is similar. But as carat increases, then price increases independent of cut 
